---
title: "Parched Peccaries; Yucatan, Mexico"
author: "jugal"
date: "02/03/2021"
---

## Drought-constrained white-lipped peccary movements in the Yucatan, Mexico

Import data shared by Dr. R. Reyna; WLP trajectories of two groups; water hole locations / binary presence 

```{r}
#set working directory to shared dropbox/me/wlp/...
setwd("C:/Users/jugal/Dropbox/Research/Movement Ecology/White lipped peccary")

#import csv of locations
peccary <- read.csv("Rafa_pecarry_13_2018_forRaja_withoutinfo.csv")

```

Clean & pre-process data: remove NAs; project; 

```{r}
#clip to keep coordinate data 
peccary <- subset(peccary, select = c("Acquisition.Time",
                                      "Acquisition.Start.Time",
                                      "GPS.Fix.Time",
                                      "GPS.Latitude",
                                      "GPS.Longitude",
                                      "GPS.UTM.Northing",
                                      "GPS.UTM.Easting",
                                      "GPS.Altitude",
                                      "GPS.Speed",
                                      "GPS.Heading"))

#missing data
library(Amelia)
missmap(peccary)

#remove all missing rows 
peccary <- peccary[complete.cases(peccary), ]
str(peccary)

#basic plot
library(tidyverse)

#long/lat
ggplot(data = peccary, mapping = aes(x = GPS.Longitude,
                                     y = GPS.Latitude)) +
  geom_point()

#clip 
peccary <- subset(peccary, GPS.Longitude >= -90.20) #fieldwork main house (locations at top left)
peccary <- subset(peccary, GPS.Longitude <= -89.70) #field house ? trip to near city?
peccary <- subset(peccary, GPS.Speed <= 20) #omit records with v greater than 20 m/s

ggplot(data = peccary, mapping = aes(x = GPS.Longitude,
                                     y = GPS.Latitude)) +
  geom_point()

#asthetic name changes
#UTM
peccary$easting <- as.numeric(peccary$GPS.UTM.Easting)
peccary$northing <- as.numeric(peccary$GPS.UTM.Northing)

#time, t
peccary$datetime <- strptime(peccary$GPS.Fix.Time,
                            format = ("%Y.%m.%d %H:%M:%S"))
peccary$datetime <- as.POSIXct(peccary$datetime)
peccary <- subset(peccary, datetime >= as.POSIXct("2017-01-01 00:00:00")) #for only 2017 data

#altitude/elevation, z
peccary$altitude <- as.numeric(peccary$GPS.Altitude)

#direction/heading
peccary$heading <- peccary$GPS.Heading

#speed
peccary$speed <- peccary$GPS.Speed

#long lat / x y
peccary$longitude <- peccary$GPS.Longitude
peccary$latitude <- peccary$GPS.Latitude

peccary <- subset(peccary, select = c("datetime",
                                      "longitude",
                                      "latitude",
                                      "altitude",
                                      "speed",
                                      "heading"))

peccary <- peccary[complete.cases(peccary), ]
str(peccary)

plot(peccary, col = "pink")

```

save
```{r}
#save_peccary <- peccary
#write.csv(save_peccary, "peccary_0522.csv")
peccary <- save_peccary
```

import waterbodies 
```{r}
agua <- read.csv("aguadas.csv")
```

sf
```{r}
library(sf)
peccary <- st_as_sf(peccary,
                    coords = c("latitude", "longitude"))
```


moveVis - core
```{r}
library(moveVis)
move_peccary <- df2move(peccary, 
                        proj = "+proj=longlat +ellp",
                        x = "longitude",
                        y = "latitude",
                        time = "datetime")

move_peccary <- align_move(move_peccary,
                           res = "mean",
                           digit = "min",
                           unit = "mins",
                           spaceMethod = "greatcircle")
```

aguadas shared by R Reyna
```{r}
#switch to local
setwd("C:/Users/jugal/Documents/research/peccaries")

#import csv of watering holes
ponds <- read.csv("aguadas0522.csv")
ponds <- subset(ponds, select = c("ID",
                           "X",
                           "Y"))

#reset working directory
setwd("C:/Users/jugal/Dropbox/Research/Movement Ecology/White lipped peccary")

library(sf)
ponds_shp <- st_read("aguadas/Waterhole_Locations.shp") 
ponds_shp <- subset(ponds_shp, select = c("ID",
                           "X",
                           "Y"))

st_geometry_type(ponds_shp)

```
```{r}
ggplot(data = ponds_shp,
       aes(x = X,
           y = Y)) +
  geom_point()
```
```{r}
ponds_shp <- st_transform(ponds_shp, crs = 4326)
```

```{r}
ponds_shp <- subset(ponds_shp, Y <= 2011647.38) 
ponds_shp <- subset(ponds_shp, X <= 214368.46 )

```

moveVis - osm - topographic
```{r}
frames <- frames_spatial(move_peccary,
                         map_service = "osm", map_type = "topographic",
                         trace_show = T,
                         trace_colour = "pink",
                         path_legend = F,
                         path_size = 0.8, 
                         path_colours = "brown") %>% 
  add_northarrow(height = 0.05,
                 position = "upperright",
                 colour = "black",
                 size = 1,label_text = "N",
                 label_margin = 0.4,
                 label_size = 5) %>%
  add_scalebar(position = "bottomleft",
               units = "km") %>%
  add_labels(x = "Longitude", y = "Latitude", 
             title = "White-lipped peccary group \n in Calakmul Biosphere Reserve",
             caption = "Peccary movements become increasingly constrained as dry season approaches") %>%
  add_timestamps(move_peccary, type = "label") %>% 
  add_gg(gg = expr(geom_sf(aes(),
                              data = ponds_shp,
                              colour = "blue")))

frames[[100]]
frames[[200]]

```






moveVis - pink grayscale
```{r}
frames <- frames_spatial(move_peccary,
                         map_service = "osm", map_type = "grayscale",
                         trace_show = T,
                         trace_colour = "pink",
                         path_legend = F) %>% 
  add_northarrow(height = 0.05,
                 position = "upperright",
                 colour = "black",
                 size = 1,label_text = "N",
                 label_margin = 0.4,
                 label_size = 5) %>%
  add_scalebar(position = "bottomleft",
               units = "km") %>%
  add_labels(x = "Longitude", y = "Latitude") %>%
  add_timestamps(move_peccary, type = "label")

frames[[100]]
frames[[200]]

```

moveVis - pink terrain
```{r}
frames <- frames_spatial(move_peccary,
                         map_service = "osm", map_type = "terrain",
                         trace_show = T,
                         trace_colour = "pink",
                         path_legend = F) %>% 
  add_northarrow(height = 0.05,
                 position = "upperright",
                 colour = "black",
                 size = 1,label_text = "N",
                 label_margin = 0.4,
                 label_size = 5) %>%
  add_scalebar(position = "bottomleft",
               units = "km") %>%
  add_labels(x = "Longitude", y = "Latitude", 
             title = "Squadron of white-lipped peccaries in the Yucatan, early 2017",
             caption = "movements become constrained as drought season approaches in Yucatan, Mexico 2017") %>%
  add_timestamps(move_peccary, type = "label")

frames[[100]]
frames[[200]]

```
moveVis - pink humanitarian
```{r}
frames <- frames_spatial(move_peccary,
                         map_service = "osm", map_type = "humanitarian",
                         trace_show = T,
                         trace_colour = "pink",
                         path_legend = F) %>% 
  add_northarrow(height = 0.05,
                 position = "upperright",
                 colour = "black",
                 size = 1,label_text = "N",
                 label_margin = 0.4,
                 label_size = 5) %>%
  add_scalebar(position = "bottomleft",
               units = "km") %>%
  add_labels(x = "Longitude", y = "Latitude") %>%
  add_timestamps(move_peccary, type = "label")

frames[[100]]
frames[[200]]

```

moveVis - pink mtb
```{r}
frames <- frames_spatial(move_peccary,
                         map_service = "osm", map_type = "mtb",
                         trace_show = T,
                         trace_colour = "pink",
                         path_legend = F) %>% 
  add_northarrow(height = 0.05,
                 position = "upperright",
                 colour = "black",
                 size = 1,label_text = "N",
                 label_margin = 0.4,
                 label_size = 5) %>%
  add_scalebar(position = "bottomleft",
               units = "km") %>%
  add_labels(x = "Longitude", y = "Latitude") %>%
  add_timestamps(move_peccary, type = "label")

frames[[100]]
frames[[200]]

```


moveVis animation functiion
```{r}
animate_frames(frames, out_file = "peccaries_osm_topographic.gif",
               overwrite = T)
```